Social Visual Analytics: DevOps for Data Scientists

### 2015-03-23 discussion.tex
various snarky comments

@@

{\bf Modularity and granularity.} The ability to partition work into
independent units (modularity) is a key to working productively in teams.
It is good if the units can be kept small, so team members can realize
benefits at least proportional to the work on the units (granularity).
RCloud's notebook and versioning capabilities allow work to be divided
into units as fine as the underlying language allows, and versioning
encourages making incremental changes at low cost and without disrupting
the work of others.

@@

{\bf Content export} is not a capability we aimed at supporting,
and R already has many packages for this. Instead, our goal was to
encourage ways of working without needing as much explicit export.
We did assume that RCloud notebooks should play well in the ecosystem
of other tools, and this worked as intended, though we learned it also
creates a temptation to absorb features almost without limit. (Should
RCloud provide templates for personal web pages and graphical
editing? Should it incorporate an email client and emacs?
Should it push results to Facebook?) Experience shows that interfaces
to external systems and other features almost always win over architectural
purity in the real world.

{\bf Social-psychological incentives} and {\bf voting and ranking}
are supported through starring. This mechanism was employed often
by our test audience. An obvious next step would be to enhance
recommendations using relationships discovered by static and
dynamic code analysis. This may be considered both within and across
collections of scripts (the latter being similar to Vistrails'
enhanced recommendations by clustering multiple workflows).
It seems valuable to know which packages are frequently used together,
or appear in proximity to a certain record types or data feeds.
In general, trivial or passive mechanisms to collect data for
recommendations are essential.

@@

{\bf Curation.} Even without formal group management, users may
curate groups of related notebooks using file system folders.
We found this particularly effective in collecting and distributing
training materials.

@@

{\bf Multiple language support.}
We did not foresee the demand to relax the assumption that programming
is done mainly in R. In retrospect, it is clear that many data science
problems are solved with cross-language approaches (for example, scikit-learn
in Python, weka in Java).\stephen{Isn't this fixed in some way now?}

@@

Our back-end environment assumes a high degree of trust between users.
Access control for information security is delegated to the host
operating system and web server. In practice, most operating
systems and web servers rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome. 
Information security in RCloud should be improved, but more
sophisticated approaches (such as information flow analysis,
and query languages that ensure differential privacy) are
active research topics, and many difficult problems remain
\cite{Moore:2011:SAF}. We are hopeful that having a richer
environment for collaboration and information sharing will encourage
new approaches to information security in visual analytics.

>>>>

\gordon{This section now overlaps with the Interview Study quite a bit - we can
  use that to motivate this, but that would mean not immediately responding to
  the complaints raised there. On the other hand, trying to merge these two
  sections would be a beast. The good news is, we anticipated most of the
  problems that people are facing now. :-/}

@@

{\bf Modularity and granularity.} The ability to partition work into
independent units (modularity) is a key to working productively in teams.
It is good if the units can be kept small, so team members can realize
benefits at least proportional to the work on the units (granularity).
RCloud's notebook and versioning capabilities allow work to be divided
into units as fine as the underlying language allows, and versioning
encourages making incremental changes at low cost and without disrupting
the work of others. \gordon{But this design choice means that developing
larger, multi-page apps is less well supported.}

@@

{\bf Content export} is not a capability we aimed at supporting,
and R already has many packages for this. Instead, our goal was to
encourage ways of working without needing as much explicit export.
We did assume that RCloud notebooks should play well in the ecosystem
of other tools, and this worked as intended, though we learned it also
creates a temptation to absorb features almost without limit. (Should
RCloud provide templates for personal web pages and graphical
editing? Should it incorporate an email client and emacs?
Should it push results to Facebook?) Experience shows that interfaces
to external systems and other features almost always win over architectural
purity in the real world. \gordon{If you don't export, doesn't that mean
you have to support every possible operation downstream? Due to popular demand,
we did add plot export.}

{\bf Social-psychological incentives} and {\bf voting and ranking} are supported
through starring. This mechanism was employed often by our test
audience. \gordon{actually, the notebook tree quickly became so unwieldy that
no one uses this.} An obvious next step would be to enhance recommendations using
relationships discovered by static and dynamic code analysis. This may be
considered both within and across collections of scripts (the latter being
similar to Vistrails' enhanced recommendations by clustering multiple
workflows).  It seems valuable to know which packages are frequently used
together, or appear in proximity to a certain record types or data feeds.  In
general, trivial or passive mechanisms to collect data for recommendations are
essential.

@@

{\bf Curation.} Even without formal group management, users may
curate groups of related notebooks using notebook tree folders.
We found this particularly effective in collecting and distributing
training materials.

@@

{\bf Multiple language support.}
We did not foresee the demand to relax the assumption that programming
is done mainly in R. In retrospect, it is clear that many data science
problems are solved with cross-language approaches (for example, scikit-learn
in Python, weka in Java).\stephen{Isn't this fixed in some way now?}
\gordon{Should we have a section on ways that we've responded to popular demand?}

@@

Our back-end environment assumes a high degree of trust between users.
Access control for information security is delegated to the host
operating system and web server. In practice, most operating
systems and web servers rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome. 
Information security in RCloud should be improved, but more
sophisticated approaches (such as information flow analysis,
and query languages that ensure differential privacy) are
active research topics, and many difficult problems remain
\cite{Moore:2011:SAF}. We are hopeful that having a richer
environment for collaboration and information sharing will encourage
new approaches to information security in visual analytics.
\gordon{This sounds like we think we are helping by pushing the
system to the breaking point, is that the idea?}
